A digital filter designing tool for the Hankel and Fourier transforms in potential-, diffusive-, and wavefield modeling

### 2017-12-29 ManuscriptGeophysics2018.tex
Well then
Make intro (1/2/3)-titles clearer

% \subsection{Review}

The introduction of linear filters to electromagnetic geophysics initiated a
wealth of investigation, publications, and development of computer programs
that extended and improved the method. A few dozen articles have been published
in the 70s and 80s by Ghosh, Koefoed, Das, Anderson, and many others, the vast
majority of it in the journal \emph{Geophysical Prospecting}.
%  (see Figure~\ref{fig:authors})
Another common name of DLF is \emph{fast Hankel transform} (FHT), popular
because of the similarity of the name to the well known \emph{fast Fourier
transform} (FFT). The name was introduced, as far as we can see, by
\cite{GP.79.Johansen} in their article of the same name. However, the name FHT
can be misleading as it has \emph{Hankel} in the name, but the DLF can also be
applied to the Fourier transform (and other linear transforms).
% \begin{figure}[tbp]
%   \centering
%   \fbox{\includegraphics[width=10cm]{figures/author-figure}}
%     \caption{List by year of publications included in the review (year and
%       first three letters of principal author).}
%     \label{fig:authors}
% \end{figure}
Publications fall broadly into one or several of three categories: (1) new
methods (application of DLF to new measurement techniques); (2) filter
improvements (new filters or providing new or improved methods for the
determination of filter coefficients); and (3) computational tools. The
following is a brief review of the most relevant publications regarding DLF,
without claiming completeness.

\subsubsection{(1) New methods}

@@

\subsubsection{(3) Codes}

@@

% \subsection{Outline}

All mentioned publications have in common that they were derived for direct
current methods (DC) or low frequency methods, such as time-domain shallow EM
methods (TEM), or controlled-source electromagnetics (CSEM), but not for high
frequency methods such as ground-pe\-ne\-tra\-ting radar (GPR). Generally it
was even thought that the filter method works only in the diffusive limit
\citep[e.g.,][]{GEO.15.Hunziker}. Also, there is no open-source filter
designing tool readily available, they are available as a described method
in articles or as code upon request by the author, if at all. The presented
algorithm \texttt{fdesign} tries to fill this gap by providing the tools to
design general or purpose-built filters using the direct matrix inversion
method. After a brief overview of the methodology, the theory, and the code we
show examples of its usage, presenting some new filters for both, CSEM and GPR
data. The algorithm and many more examples of its usage can be found on
\href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.
The examples can be used as templates to design new filters.

@@

We except the presented algorithm is useful for at least 3 scenarios:
\begin{enumerate}
  \item Provide a fast method to design problem-specific filters. For bigger
    inversion projects (such as, for instance, massive stochastic CSEM
    inversions) a purpose designed, short filter might save much more time than
    it costs to design it. This could even be integrated into inversion codes,
    as an optional pre-inversion step.
  \item Extend the filter method to new areas, namely higher frequencies and
    acoustic and elastic wavefields.
  \item Raise a new interest for digital linear filters in geophysics, by
    making it very easy for anyone to play around, create their own filters,
    and get a better understanding of it. We are sure there are many great
    filters to be discovered.
\end{enumerate}

\section{Acknowledgment}
D.W. would like to thank his colleague Oleg Gurevich Titov, who introduced him
to linear digital filters in the first place, and Jürg Hunziker for valuable
feedback which improved this manuscript considerably.
% This project would not have been successful without the fruitful discussions
% with Evert Slob and Kerry Key. Many thanks as well to Kerry Key for providing
% me with his Matlab scripts, which served as a starting point and initiated
% this project.

>>>>

The introduction of linear filters to electromagnetic geophysics initiated a
wealth of investigation, publications, and development of computer programs
that extended and improved the method. A few dozen articles have been published
in the 70s and 80s by Ghosh, Koefoed, Das, Anderson, and many others, the vast
majority of it in the journal \emph{Geophysical Prospecting}. Another common
name of DLF is \emph{fast Hankel transform} (FHT), popular because of the
similarity of the name to the well known \emph{fast Fourier transform} (FFT).
The name was introduced, as far as we can see, by \cite{GP.79.Johansen} in
their article of the same name. However, the name FHT can be misleading as it
has \emph{Hankel} in the name, but the DLF can also be applied to the Fourier
transform (and other linear transforms). Publications fall broadly into one or
several of three categories: (1) new applications (extend the usage of DLF to
other measurement techniques); (2) filter improvements (new filters or
providing new or improved methods for the determination of filter
coefficients); and (3) computational tools. The following is a brief review of
the most relevant publications regarding DLF, without claiming completeness.

\subsubsection{(1) New applications}

@@

\subsubsection{(3) Computational tools}

@@

All mentioned publications have in common that they were derived for direct
current methods (DC) or low frequency methods, such as time-domain shallow EM
methods (TEM), or controlled-source electromagnetics (CSEM), but not for high
frequency methods such as ground-pe\-ne\-tra\-ting radar (GPR). Generally it
was even thought that the filter method works only in the diffusive limit
\citep[e.g.,][]{GEO.15.Hunziker}. Also, there is no open-source filter
designing tool readily available, they are available as a described method
in articles or as code upon request by the author, if at all. The presented
algorithm \texttt{fdesign} tries to fill this gap by providing the tools to
design general or purpose-built filters using the direct matrix inversion
method. The current article falls therefore into all three mentioned categories
by applying DLF also to wave phenomena and providing new filters with the
accompanying code. After a brief overview of the methodology, the theory, and
the code we show examples of its usage, presenting some new filters for both,
CSEM and GPR data. The algorithm and many more examples of its usage can be
found on
\href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.
The examples can be used as templates to design new filters.

@@

We except the presented algorithm is useful for at least 3 scenarios:
\begin{enumerate}
  \item Provide a fast method to design problem-specific filters. For bigger
    inversion projects (such as, for instance, massive stochastic CSEM
    inversions) a purpose designed, short filter might save much more time than
    it costs to design it. This could even be integrated into inversion codes,
    as an optional pre-inversion step.
  \item Extend the filter method to new areas, namely higher frequencies and
    acoustic and elastic wavefields.
  \item Raise a new interest for digital linear filters in geophysics, by
    making it very easy for anyone to play around, create their own filters,
    and get a better understanding of it. We are sure there are many great
    filters to be discovered.
\end{enumerate}
~
